### 2020-11-05 meetingquestions.Rmd
added meeting q's

---
title: "meeting q"
author: "Joe Despres"
date: "11/5/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(patchwork)
theme_set(theme_light())
data <- read_csv(here::here("data", "df.csv"))
```

```{r}
dep_vars <- select(data, ends_with("_fh")) %>% names()
resp_vars <- select(data, -ends_with("_fh")) %>% names()
dep_vars
resp_vars
```

Our dependent variables are indexes measured in increments of 0.5, can we use regression methods without making extreme transformations?

```{r}
p1 <- data %>% 
  ggplot(mapping = aes(x = -freedix_fh)) + 
  geom_histogram(bins = 15) + labs(title = "Freedom Index", subtitle = 'Further Right More "Free"')

p2 <- data %>% 
  ggplot(mapping = aes(x = -polrights_fh)) + 
  geom_histogram(bins = 15) + labs(title = "Political Rights Index")

p3 <- data %>% 
  ggplot(mapping = aes(x = -cliberities_fh)) + 
  geom_histogram(bins = 15) + labs(title = "Civil Liberties Index")

p1 + p2 + p3

```

Missing values

```{r}
skimr::skim(data)
str(data)
data %>% view()
```

Best subsets resuls show most vars are insignificant. Can we procede with the project?

```{r}
models <- list()
for (i in 1:length(vars)) {
  vc <- combn(vars,i)
  for (j in 1:ncol(vc)) {
    model <- as.formula(paste0(outcome, " ~", paste0(vc[,j], collapse = " + ")))
    models <- c(models, model)
  }
}

subsets <- map(models, function(x) lm(x, data)) %>% 
  map(broom::glance) %>% 
  setNames(models) %>% 
  bind_rows(.id = "id") %>% 
  distinct() %>% 
  rename(model = id)

```

```{r}
r <- subsets %>% group_by(df) %>% summarise(adj = max(adj.r.squared, na.rm = T))

subsets %>% 
  ggplot(aes(x = df, y = adj.r.squared)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, adj), se = FALSE, span = 0.5, color = "Green") + 
  labs(title = "Best Subsets", x = "Parameters", y = "Adjusted R Squared")
```
at 5 degrees of freedom
```{r}
model_df5 <- subsets %>% filter(df == 5) %>% 
  arrange(desc(adj.r.squared)) %>% 
  select(model) %>% head(1) %>% as.character()

model_df5
fit <- lm(model_df5, data = data)
summary(fit)
```

```{r}
hist(resid(fit))
```

at 4 degrees of freedom
```{r}
model_df4 <- subsets %>% filter(df == 4) %>% 
  arrange(desc(adj.r.squared)) %>% 
  select(model) %>% head(1) %>% as.character()

fit <- lm(model_df4, data = data)
summary(fit)
```