### 2020-11-11 isoslides.Rmd
.

---
title: "slides"
author: "Joe Despres"
date: "11/9/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
source(here::here("R", "regression.R"))
options(htmltools.dir.version = FALSE, echo = FALSE)
```

# Introduction
- political freedom as a function of 
- 
---
# Model
```{r}
equatiomatic::extract_eq(model = model_df5, intercept = "beta", terms_per_line = 1)
```

$$
\operatorname{polrights\_fh} = \beta_{0} + 
\beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) + \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +
\beta_{3}(\operatorname{life\_expectancy\_years}) + \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) + \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
$$

---
```{r echo=FALSE}
subsets %>% 
  mutate_if(is.numeric, round, 3) %>% 
  select(model, adj.r.squared) %>% 
DT::datatable(fillContainer = FALSE, options = list(pageLength = 6))
```
---

# Model

# Findings

# Conclusions